---
title: "Data Mining"
author: "Orange 9"
date: "2022-10-02"
output: html_document
---

```{r setup}
library(tidyverse)
library(ggplot2)
library(arules)
library(arulesViz)
```

## Data Mining Homeworks

```{r data}
df_mws<-read.csv("https://raw.githubusercontent.com/sjsimmo2/DataMining-Fall/master/orderData.csv")
```

## Report 1

Eats4Life would like to update its menu to include wine suggestions with each of its main entrees (defined by the meat selection). The owner would like to take a Data Analytics approach and explore data he collected over the past several years on main courses (meat) and wine that was ordered with it. Eats4Life is open to listing more than one wine for each main entree, but only if the data supports it. The scope of services requested include:

```{r prep}
# create unique ID by combining order and seat
df_mws$ID <- paste(df_mws$orderNo, df_mws$seatNo, sep="_")

# filter out sides
df_mw <- df_mws %>% filter(row_number() %% 3 != 0)

#convert long to wide
trans.dat <- as(split(df_mw$item, df_mw$ID), "transactions")

# take a look at transaction data
#inspect(trans.dat)
trans.dat@itemInfo$labels

# what are the top 5 items?
itemFrequencyPlot(trans.dat,topN=5,type="relative")
```

-   Summary information on the main entrees (meat)

    ```{r}
# list of meats - 8 options
df_m <- df_mws[seq(1, NROW(df_mws), by = 3),]
meats <- unique(df_m$item)

#create empty data frame 
mydata <- data.frame(matrix(ncol = 8, nrow = 0))

# Looking for if any combinations of meat/wine doesn't exist
# loop through list of meats as lhs
for (i in 1:length(meats)){
  mydata <- rbind(mydata, inspect(apriori(trans.dat, parameter = list(supp=0.0, conf=0.0),appearance = list(lhs=meats[i], default="rhs"), minlen=2)))
}
colnames(mydata) <- c('lhs', 'arrow', 'rhs', 'support', 'confidence', 'coverage', 'lift', 'count')
View(mydata)

# see if any supports are 0
mydata %>%
  arrange(mydata$lhs, mydata$support) %>%
  View(.)

# list of wines - 13 options
df_w <- df_mws[seq(2, NROW(df_mws), by = 3),]
wines <- unique(df_w$item)
table(df_w$item)

    ```
```{r}
# look at data with order number
df_mw %>%
  group_by(orderNo) %>%
  count(item) %>%
  arrange(desc(n)) %>%
  ungroup()

# see if seatNo affects order of meat
seat1 <- df_m %>%
  filter(seatNo == 1) %>%
  count(item)
seat1$percent = seat1$n/sum(seat1$n) * 100
seat2 <- df_m %>%
  filter(seatNo == 2) %>%
  count(item)
seat2$percent = seat2$n/sum(seat2$n) * 100
seat3 <- df_m %>%
  filter(seatNo == 3) %>%
  count(item)
seat3$percent = seat3$n/sum(seat3$n) * 100
seat4 <- df_m %>%
  filter(seatNo == 4) %>%
  count(item)
seat4$percent = seat4$n/sum(seat4$n) * 100
seat5 <- df_m %>%
  filter(seatNo == 5) %>%
  count(item)
seat5$percent = seat5$n/sum(seat5$n) * 100

# run Pearson Chi Sq Test
chisq.test(c(seat1$percent, seat2$percent, seat3$percent, seat4$percent, seat5$percent))
```

-   Wine suggestion(s) for each main entree along with supporting information as to why this (these) wines are suggested for the entrée (if you have no suggested wine for a given entrée, provide information as to why this is your suggestion)
```{r}
# see if seatNo affects order of wine
seat1.w <- df_w %>%
  filter(seatNo == 1) %>%
  count(item)
seat1.w$percent = seat1.w$n/sum(seat1.w$n) * 100
seat2.w <- df_w %>%
  filter(seatNo == 2) %>%
  count(item)
seat2.w$percent = seat2.w$n/sum(seat2.w$n) * 100
seat3.w <- df_w %>%
  filter(seatNo == 3) %>%
  count(item)
seat3.w$percent = seat3.w$n/sum(seat3.w$n) * 100
seat4.w <- df_w %>%
  filter(seatNo == 4) %>%
  count(item)
seat4.w$percent = seat4.w$n/sum(seat4.w$n) * 100
seat5.w <- df_w %>%
  filter(seatNo == 5) %>%
  count(item)
seat5.w$percent = seat5.w$n/sum(seat5.w$n) * 100

# run Pearson Chi Sq Test
chisq.test(c(seat1.w$percent, seat2.w$percent, seat3.w$percent, seat4.w$percent, seat5.w$percent))

```

```{r associations}
# get rules
rules <- apriori(trans.dat, parameter = list(supp = 0.001, conf = 0.001, target="rules"), minlen=2)
rules<-sort(rules, by="confidence", decreasing=TRUE)

inspect(rules[1:11])

plot(rules)

top10rules = head(rules, n = 10, by = "confidence")
plot(top10rules, method = "graph",  engine = "htmlwidget")
```

```{r meatrules}
# loop through list of meats as lhs
for (i in 1:length(meats)){
  assign(paste0(meats[i], ".rule1"), apriori(trans.dat, parameter = list(supp=0.001, conf=0.001),appearance = list(lhs=meats[i], default="rhs"), minlen=2))
}

# loop through list of meats as rhs
for (i in 1:length(meats)){
  assign(paste0(meats[i], ".rule2"), apriori(trans.dat, parameter = list(supp=0.001, conf=0.001),appearance = list(rhs=meats[i], default="lhs"), minlen=2))
}

# sort for ease
Swordfish.rule1<-sort(Swordfish.rule1, by="confidence", decreasing=TRUE)
Swordfish.rule2<-sort(Swordfish.rule2, by="confidence", decreasing=TRUE)

Salmon.rule1<-sort(Salmon.rule1, by="confidence", decreasing=TRUE)
Salmon.rule2<-sort(Salmon.rule2, by="confidence", decreasing=TRUE)

```

-   Any other information of interest in terms of customer order habits

```{r habits}
# number of parties = 24687
length(unique(df_mws$orderNo))
# largest party size = 5
unique(df_mws$seatNo)
# composition of parties
table(df_mws$seatNo)
  # 12,759 parties of 5
  # 19,368 parties of 4
  # 6,837 parties of 3
  # 31,824 parties of 2
  # 3,273 parties of 1

#create empty data frame 
df <- data.frame(matrix(ncol = 8, nrow = 0))

#Looking at highest lifts and supports for each entree
# loop through list of meats as lhs
for (i in 1:length(meats)){
  df <- rbind(df, inspect(apriori(trans.dat, parameter = list(supp=0.001, conf=0.001),appearance = list(lhs=meats[i], default="rhs"), minlen=2)))
}
colnames(df) <- c('lhs', 'arrow', 'rhs', 'support', 'confidence', 'coverage', 'lift', 'count')
View(df)

df %>%
  arrange(df$lhs, desc(df$lift)) %>%
  View(.)

df %>%
  arrange(df$lhs, desc(df$support)) %>%
  View(.)
```
```{r}
# creating visuals for wine
# dataset with wines
wine.df <- df_mws[c(FALSE, TRUE, FALSE), ]
aggr_wine <- wine.df %>% 
  group_by(item) %>% 
  summarize(total_cnt = n())

View(aggr_wine)

# creating the percentage column 
aggr_wine$percentage_of_wines = round(aggr_wine$total_cnt / sum(aggr_wine$total_cnt) * 100, 2)

# arranging the dataset
aggr_wine <- aggr_wine %>% 
  arrange(percentage_of_wines)


# plotting the percentages of wine
plot <- ggplot(aggr_wine, aes(x=reorder(item, -percentage_of_wines), percentage_of_wines)) + geom_bar(stat='identity') + 
  geom_text(aes(label=percentage_of_wines), position = position_dodge(0.9), vjust=1, colour='white') + 
  labs(title = 'Distribution of the Wines', x = 'Wine Type', 
       y = 'Percentage of Orders') + theme(axis.text.x = element_text(angle = 90))
plot
```
```{r}
# visualizing entrees 
#dataset with only entrees
entrees.df <- df_mws[c(TRUE,rep(FALSE,2)), ]

# creating an aggregate 
aggr_entrees <- entrees.df %>% 
  group_by(item) %>% 
  summarize(total_cnt = n())

View(aggr_entrees)

# creating the percentage column 
aggr_entrees$percentage_of_entrees = round(aggr_entrees$total_cnt / sum(aggr_entrees$total_cnt) * 100, 2)

plot.entrees <- ggplot(aggr_entrees, aes(x=reorder(item, -percentage_of_entrees), percentage_of_entrees)) + geom_bar(stat='identity') + 
  geom_text(aes(label=percentage_of_entrees), position = position_dodge(0.9), vjust=1, colour='white') + 
  labs(title = 'Distribution of the Entrees', x = 'Entree Type', 
       y = 'Percentage of Orders') + theme(axis.text.x = element_text(angle = 90))
plot.entrees
```
```{r}
#visualizing the count of average number of people per table and plot distribution of the number of people per table 
aggr_order_no <- df_mws %>% 
  group_by(orderNo) %>%
  summarize(max_num = max(seatNo))

mean(aggr_order_no$max_num)
median(aggr_order_no$max_num)
seats_dist <- aggr_order_no %>% 
  group_by(max_num) %>% 
  summarize(dist_of_seats = n()) 
data.frame(seats_dist)
seats_dist <- seats_dist[order(seats_dist$dist_of_seats),]
seats_dist$percentage <- round(seats_dist$dist_of_seats / sum(seats_dist$dist_of_seats) * 100, 2) 


plot <- ggplot(seats_dist, aes(max_num, percentage)) + geom_bar(stat='identity') + 
  geom_text(aes(label=percentage), position = position_dodge(0.9), vjust=1, colour='white') + 
  labs(title = 'Distribution of the Number of Customers per Table', x = 'Number of Customers per Table', 
       y = 'Percentage of Meals')
plot
```

